---
title: "Classification and Regression Algorithms On Online News Popularity Dataset"
author: "Brendan Dagys"
date: "9/21/2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PREPROCESSING

```{r, message = FALSE}

# install.packages('caret')
library(caret)
library(MASS)
setwd('/')

news = read.csv('/Users/brendan/Desktop/Personal R Projects/Online News Popularity/OnlineNewsPopularity.csv', header = TRUE)
```

Preliminary check of the data.

```{r}
summary(news)
str(news)
sum(is.na(news))
```

Filter out 1 outlier with high values in 'n_unique_tokens', 'n_non_stop_words', and 'n_non_stop_unique_tokens'. Deduced from the summary function.

```{r}
news = news[!news$n_unique_tokens == 701,]
```

Remove non-predictive variables.

```{r}
news = subset(news, select = -c(url, timedelta, is_weekend))
```

Convert categorical variables into factors with 2 levels.

```{r}
news$weekday_is_monday = factor(news$weekday_is_monday) 
news$weekday_is_tuesday = factor(news$weekday_is_tuesday) 
news$weekday_is_wednesday = factor(news$weekday_is_wednesday) 
news$weekday_is_thursday = factor(news$weekday_is_thursday) 
news$weekday_is_friday = factor(news$weekday_is_friday) 
news$weekday_is_saturday = factor(news$weekday_is_saturday) 
news$weekday_is_sunday = factor(news$weekday_is_sunday) 

news$data_channel_is_lifestyle = factor(news$data_channel_is_lifestyle) 
news$data_channel_is_entertainment = factor(news$data_channel_is_entertainment) 
news$data_channel_is_bus = factor(news$data_channel_is_bus) 
news$data_channel_is_socmed = factor(news$data_channel_is_socmed) 
news$data_channel_is_tech = factor(news$data_channel_is_tech) 
news$data_channel_is_world = factor(news$data_channel_is_world)
```

Plotting a histogram of the dependent variable shows a large right skew.

```{r}
hist(news$shares, xlim = c(0, 50000), ylim = c(0, 40000), breaks = 100)
```

I create a new table with the dependent variable changed by a log-transform.

```{r}
news_log = news
news_log$shares = log(news_log$shares)
```

Let's plot a histogram.

```{r}
hist(log(news$shares))
```

This looks much better and resembles a normal distribution!

Create a copy with a binary 'share' variable for > 1400 shares.

```{r}
news_binary = news
news_binary$shares = as.factor(ifelse(news_binary$shares > 1400, 'high', 'low'))
```

# SAMPLING AND PARTITIONING

Sample the dataset into 70% training and 30% test data.

```{r}
train_index = sample(nrow(news), as.integer(nrow(news) * 0.7))
```

Alternatively, we can use the 'caret' package to do the same sampling.

```{r}
# train_index = createDataPartition(y = news$shares, p = 0.7, list = FALSE)
```

Here, I generate training and test sets using the index.

```{r}
news_train = news[train_index,]
news_test = news[-train_index,]

news_log_train = news_log[train_index,]
news_log_test = news_log[-train_index,]

news_binary_train = news_binary[train_index,]
news_binary_test = news_binary[-train_index,]
```

# LINEAR REGRESSION

Fit a model using all variables, with 'shares' being the dependent variable.

```{r}
set.seed(7)
linear_model = lm(shares ~ ., data = news_train)
summary(linear_model)
```

R-squared value is 0.02187 and adjusted is 0.01993, which are quite low.

Calculating the RMSE gives us a value of 9679.06:

```{r, warning = FALSE}
predict_linear_model = predict(linear_model, news_test)
sqrt(mean((predict_linear_model - news_test$shares)^2))
```

Let's try another linear regression using the log-transformed values.

```{r}
linear_model_log = lm(shares ~ ., data = news_log_train)
summary(linear_model_log)
```

We now have a higher R-squared value of 0.1269 and adjusted: 0.1251
The model can explain 12% of the variation of the 'shares' variable!

This time the RMSE value is 0.8755.
```{r}
predict_linear_model_log = predict(linear_model_log, news_log_test)
sqrt(mean((predict_linear_model_log - news_log_test$shares)^2))
```

# FEATURE SELECTION

Using the p-values of the regression output to feature select:

```{r}
news_binary_train3 = news_binary_train[, c('num_hrefs', 'num_self_hrefs', 'num_imgs', 'average_token_length', 'num_keywords', 'data_channel_is_lifestyle', 'data_channel_is_entertainment', 'data_channel_is_bus', 'data_channel_is_socmed', 'data_channel_is_tech', 'kw_min_min', 'kw_min_max', 'kw_avg_max', 'kw_avg_max', 'kw_min_avg', 'kw_max_avg', 'kw_avg_avg', 'weekday_is_monday', 'weekday_is_tuesday', 'weekday_is_wednesday', 'weekday_is_thursday', 'weekday_is_friday', 'LDA_00', 'LDA_02', 'global_subjectivity', 'title_sentiment_polarity', 'abs_title_subjectivity', 'shares')]

news_binary_test3 = news_binary_test[, c('num_hrefs', 'num_self_hrefs', 'num_imgs', 'average_token_length', 'num_keywords', 'data_channel_is_lifestyle', 'data_channel_is_entertainment', 'data_channel_is_bus', 'data_channel_is_socmed', 'data_channel_is_tech', 'kw_min_min', 'kw_min_max', 'kw_avg_max', 'kw_avg_max', 'kw_min_avg', 'kw_max_avg', 'kw_avg_avg', 'weekday_is_monday', 'weekday_is_tuesday', 'weekday_is_wednesday', 'weekday_is_thursday', 'weekday_is_friday', 'LDA_00', 'LDA_02', 'global_subjectivity', 'title_sentiment_polarity', 'abs_title_subjectivity', 'shares')]
```

Using the varImp function from caret to do some feature selection, and saving the top-15 into the array new_variables:

```{r}
importance = varImp(linear_model)
importance[order(importance, decreasing = 1), , drop = F]
(new_variables = rownames(importance[order(importance, decreasing = 1), , drop = F])[1:15])

news_binary_train2 = news_binary_train[, c('kw_avg_avg', 'kw_max_avg', 'data_channel_is_entertainment', 'kw_min_avg', 'num_hrefs', 'data_channel_is_lifestyle', 'abs_title_subjectivity', 'self_reference_min_shares', 'average_token_length', 'data_channel_is_tech', 'data_channel_is_bus', 'data_channel_is_world', 'global_subjectivity', 'data_channel_is_socmed', 'n_tokens_title', 'num_self_hrefs', 'kw_min_min', 'global_rate_positive_words', 'n_tokens_content', 'self_reference_max_shares', 'shares')]

news_binary_test2 = news_binary_test[, c('kw_avg_avg', 'kw_max_avg', 'data_channel_is_entertainment', 'kw_min_avg', 'num_hrefs', 'data_channel_is_lifestyle', 'abs_title_subjectivity', 'self_reference_min_shares', 'average_token_length', 'data_channel_is_tech', 'data_channel_is_bus', 'data_channel_is_world', 'global_subjectivity', 'data_channel_is_socmed', 'n_tokens_title', 'num_self_hrefs', 'kw_min_min', 'global_rate_positive_words', 'n_tokens_content', 'self_reference_max_shares', 'shares')]
```

Let's include only statistically significant variables. I'll use the Stepwise regression function.
The Akaike Information Criterion is an estimator of the relative quality of statistical models for a given
set of data. AIC estimates the quality of each model, relative to each of the other models. Thus, AIC
provides a means for feature selection. AIC estimates the relative information lost when a given model is used
to represent the process that generated the data. It does not provide absolute quality; only the quality relative
to other data models. Given a set of candidate models for the data, the preferred model is the one with the
minimum AIC value.

Running the function backward yields a lower AIC than running forward (7143 vs. 7819)

Backward is better!

```{r, include = FALSE}
null <- lm(shares ~ 1, data = news_log_train)
full <- lm(shares ~ ., data = news_log_train)

stepF <- stepAIC(null, scope = list(lower = null, upper = full), direction = 'forward', trace = TRUE)
stepB <- stepAIC(full, direction = 'backward', trace = TRUE)
linear_model_log_step = stepAIC(linear_model_log)
```

The code looks as follows. Because of the extremely long output, I will not print in the R Markdown file.

null <- lm(shares ~ 1, data = news_log_train) \n
full <- lm(shares ~ ., data = news_log_train)

stepF <- stepAIC(null, scope = list(lower = null, upper = full), direction = 'forward', trace = TRUE) \n
stepB <- stepAIC(full, direction = 'backward', trace = TRUE)

```{r}
summary(stepF)
summary(stepB)
```

You can run the function this way to compute both at the same time!

linear_model_log_step = stepAIC(linear_model_log)

Only statistically significant variables remain:

```{r}
summary(linear_model_log_step)
```

Residual standard error: 0.8679
R-squared: 0.1267
Adjusted R-squared: 0.1254

```{r}
par(mar = c(2, 2, 2, 2))
par(mfrow = c(2, 2))
plot(linear_model_log_step)
```

The first plot is the residual vs. fitted plot.

# DECISION TREE

```{r, message = FALSE}
# install.packages('party')
library(party)

news_binary_tree_model = ctree(shares ~ ., data = news_binary_train)
news_binary_tree_pred = predict(news_binary_tree_model, newdata = news_binary_test)
confusionMatrix(news_binary_test$shares, news_binary_tree_pred)
```

We get an accuracy of 0.6472

Let's try another package!

```{r}
# install.packages('rpart')
library(rpart)

tree_model = rpart(shares ~ ., method = 'class', data = news_binary_train)
tree_model_pred = predict(tree_model, method = 'class', newdata = news_binary_test)
```

The resulting prediction is a matrix with two columns. We must transform it into factor form and remove the second column.

```{r}
tree_model_pred[, 1] = sapply(tree_model_pred[, 1], function(x) ifelse(x >= 0.5, 'high', 'low'))
tree_model_pred = as.factor(tree_model_pred[, 1])
confusionMatrix(news_binary_test$shares, tree_model_pred)
```

We obtain a lower accuracy.

```{r}
summary(tree_model)
```

Display cp table.

```{r}
printcp(tree_model)
```

If we wanted to create a postscript plot of the tree:

```{r}
# post(tree_model, file = filepath)
```

Now we will prune the tree to try and get a higher accuracy and reduce overfitting
Typically, you will want to select a tree size that minimizes the cross-validated error,
the xerror column printed by printcp( ). Use printcp( ) to examine the cross-validated error results, 
select the complexity parameter associated with minimum error, and place it into the prune( ) function.

```{r}
prune(tree_model, cp = tree_model$cptable[which.min(tree_model$cptable[,"xerror"]),"CP"])
tree_model_pred = predict(tree_model, method = 'class', newdata = news_binary_test)
tree_model_pred[, 1] = sapply(tree_model_pred[, 1], function(x) ifelse(x >=0.5, 'high', 'low'))
tree_model_pred = as.factor(tree_model_pred[, 1])
confusionMatrix(news_binary_test$shares, tree_model_pred)

# plot(tree_model); text(tree_model)
```

This actually gave us the same accuracy. Perhaps pruning was not necessary in this instance.

# Let's try regression trees to predict the continuous variable 'shares' from the 'news' dataframe

```{r}
news_tree_model = ctree(shares ~ ., data = news_train)
news_tree_pred = predict(news_tree_model, newdata = news_test)

summary(news_tree_pred)
```

3468 vs. 3365

```{r}
mean(news_test$shares); mean(news_tree_pred)
```

The RMSE is 103.4774

```{r}
sqrt(mean(news_test$shares - news_tree_pred[, 1])^2)
```

Using 'rpart' package this time.

```{r}
tree_model2 = rpart(shares ~ ., method = 'anova', data = news_train)
tree_model_pred2 = predict(tree_model2, method = 'anova', newdata = news_test)
```

RMSE is 49.084. Second run, it is 282.949. It varies as the training index is different each iteration.

```{r}
sqrt(mean(news_test$shares - tree_model_pred2)^2)
```

3566 vs. 3283

```{r}
mean(news_test$shares); mean(tree_model_pred2)
```

# NAIVE BAYES

```{r}
# install.packages('e1071')
library(e1071)
library(caret)

naive_model = naiveBayes(shares ~ ., data = news_binary_train)
naive_model

naive_prediction = predict(naive_model, news_binary_test)
caret::confusionMatrix(news_binary_test$shares, naive_prediction)
```

This gives us an accuracy of 0.5915

Let's try Naive Bayes again, but this time training the model.

```{r, warning = FALSE}
naive_model_trained = train(shares ~ ., data = news_binary_train, method = "nb", trControl = trainControl(method = 'cv', number = 10))
naive_model_trained

naive_trained_prediction = predict(naive_model_trained, news_binary_test)
caret::confusionMatrix(news_binary_test$shares, naive_trained_prediction)
```

This time we get an accuracy of 0.6116

# kNN

```{r}
library(class)

num_check = sapply(news, is.numeric)
news_numeric_train = news_binary_train[, num_check]
news_numeric_test = news_binary_test[, num_check]

news_knn = knn(news_numeric_train[-45], news_numeric_test[-45], news_numeric_train$shares, 10)
```

We get an accuracy of 0.5656

```{r}
caret::confusionMatrix(news_numeric_test$shares, news_knn)
```

# LOGISTIC REGRESSION

```{r}
news_binary_train2 = news_binary_train
news_binary_test2 = news_binary_test

news_binary_train2[, 58] = sapply(news_binary_train2[, 58], function(x) ifelse(x == 'low', 0, 1))
news_binary_test2[, 58] = sapply(news_binary_test2[, 58], function(x) ifelse(x == 'low', 0, 1))

logistic_model = glm(shares ~ ., data = news_binary_train2, family = binomial)
logistic_pred = predict(logistic_model, news_binary_test2, type = 'response')
logistic_pred = ifelse(logistic_pred >= 0.5, 'high', 'low')
```

Accuracy is 0.6515

```{r}
caret::confusionMatrix(news_binary_test$shares, as.factor(logistic_pred))
```

# TRAIN FUNCTION EXPLORATION

We will train the following: Naive Bayes, lm

Accuracy = 0.6101

```{r, message = FALSE}
# install.packages('naivebayes')
library(naivebayes)

train_model = train(shares ~ ., data = news_binary_train, method = 'naive_bayes')
train_pred = predict(train_model, news_binary_test)
caret::confusionMatrix(news_binary_test$shares, train_pred)
```

RMSE = 0.8725 (0.8655 from my function), R-squared = 0.1297, adjusted R-squared = 0.128

```{r, warning = FALSE}
train_model2 = train(shares ~ ., data = news_log_train, method = 'lm')
summary(train_model2)
train_pred2 = predict(train_model2, news_log_test)
sqrt(mean((train_pred2 - news_log_test$shares)^2))
```

# RANDOM FOREST

```{r, message = FALSE}
# install.packages('randomForest')
library(randomForest)

forest_model = randomForest(shares ~ ., data = news_binary_train, ntree = 200)
forest_model_pred = predict(forest_model, news_binary_test)

# The random forest gives us an accuracy of 0.6612!
caret::confusionMatrix(news_binary_test$shares, forest_model_pred)
```

# RANDOM FOREST WITH FEATURE SELECTION FROM VARIMP

```{r}
forest_model = randomForest(shares ~ ., data = news_binary_train2, ntree = 200)
forest_model_pred = predict(forest_model, news_binary_test2)

# Using feature selection lowered our accuracy slightly.
caret::confusionMatrix(news_binary_test2$shares, forest_model_pred)
```

# RANDOM FOREST WITH FEATURE SELECTION FROM REGRESSION P-VALUES

```{r}
forest_model = randomForest(shares ~ ., data = news_binary_train3, ntree = 200)
forest_model_pred = predict(forest_model, news_binary_test3)

# Using feature selection lowered our accuracy slightly.
caret::confusionMatrix(news_binary_test3$shares, forest_model_pred)
```

# CROSS VALIDATION

```{r}
train_control = trainControl('cv', 10)
train_model4 = train(shares ~ ., data = news_binary, method = 'rf', trControl = train_control)
confusionMatrix(train_model4)
```

# QUESTION ANSWERING

```{r}
# Does posting an article on a certain day of the week increase its chance of being shared on social media?

# 3903
mean(news$shares[news$weekday_is_saturday == 1 | news$weekday_is_sunday == 1])

# 3319
mean(news$shares[news$weekday_is_saturday == 0 & news$weekday_is_sunday == 0])

# Does a lengthy title reduce an article’s chance of being shared on social media?

# 0.09
cor(news$n_tokens_title, news$shares)
plot(news$n_tokens_title, news$shares)

# What genre of article (i.e. business, lifestyle, sports) is most likely to be shared on social media?
mean(news$shares[news$data_channel_is_bus])
mean(news$shares[news$data_channel_is_entertainment])
mean(news$shares[news$data_channel_is_lifestyle])
mean(news$shares[news$data_channel_is_socmed])
mean(news$shares[news$data_channel_is_tech])
mean(news$shares[news$data_channel_is_world])
```
